---
title: "Practical Machine Learning Course Project Write Up"
author: "Scot Shields"
date: "Wednesday, April 22, 2015"
output: html_document
---

##Goal:  
  
Apply machine learning techniques to collected activity data in order to predict how participants did exercises.  
  
##PreProcessing:
  
###Load Libraries:
```{r}
library(caret)
library(rpart)
library(randomForest)
```  
  
###Load Data:  
```{r}
setwd("~//Coursera DS//Machine/Project")
if(!exists("trainData"))
{
trainData <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
}

if(!exists("testData"))
{
testData <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
}
```  
###Clean Data:  
  
Remove variables with many NA or missing observations.  
```{r}  
trainClean <- trainData[, colSums(is.na(trainData)) == 0] 
testClean <- testData[, colSums(is.na(testData)) == 0]
```  
  
Find and Remove variables with little variance.
```{r}
trainDataLV <- nearZeroVar(trainClean, saveMetrics=TRUE)
trainDataLV[trainDataLowVar$nzv=="TRUE",]

trainFinal <- subset(trainClean, select = -c(X,new_window) )
testFinal <- subset(testClean, select = -c(X,new_window) )
```  
  
###Partitioning Data for Training/Model Building  
  
```{r}
set.seed(13518) # For reproducibile purpose
inTrain <- createDataPartition(trainFinal$classe, p=0.70, list=F)
train <- trainFinal[inTrain, ]
test_train <- trainFinal[-inTrain, ]
```  
  
###Training Random Forrest Model  
  
```{r}
if(!exists("model1"))
{
model1 <- train(classe ~ ., data=train, method="rf", trControl=trainControl(method="cv", 5), ntree=251)
}
```  
  
###Estimating Model Performance  
  
```{r}
PredictMod1 <- predict(model1, test_train)
confusionMatrix(test_train$classe, PredictMod1)

OutOfSampleError <- 1 - as.numeric(confusionMatrix(test_train$classe, PredictMod1)$overall[1])
OutOfSampleError
```  
From Cross Validation we see the that the model is 99.98% accurate and the out of sample error is 0.017%.
  
###Model Predictions on Test Data  
  
```{r}
result<-predict(model1,testFinal)
result
```


